---
title: "Hypothesis 1"
format:
  html:
    toc: true
    html-math-method: katex
    css: styles.css
    code-fold: true
    code-summary: "Show the code"
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r, message= FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(dplyr)
library(data.table)
library(lme4)
library(ggplot2)
library(pscl) # for odTest
library(arm)
library(sjPlot)
library(rstatix)
library(car)
library(glue)

remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}

```

```{r, include=FALSE}
wd <- getwd()
system('7z x data/assignments.zip -odata/.')
assignments <- fread("data/assignments.csv") %>% 
  mutate(course_item_type_desc = factor(course_item_type_desc), 
         vis_b = ifelse(vis_tw==0, 0, 1), 
         vis_before_attempt_tw_b = ifelse(vis_before_attempt_tw==0, 0, 1),
         vis_before_attempt_tw_b = factor(vis_before_attempt_tw_b)) %>% 
  group_by(course_item_id) %>% 
  mutate(difficulty =  sum(first_attempt == 0) / n()) %>% 
  dplyr::rename(CFA = first_attempt, forum_attendance_tw = vis_before_attempt_tw_b, forum_visits_tw = vis_before_attempt_tw) %>% 
  ungroup() 

#tw <- assignments %>% filter(assignment_time < 116) # delete observations over 3d quantile
```

## Research questions

  1. What is the relationship between Correctness of the first attempt and forum attendance during task performance? 
  
## Database desciption

  - **assignments** - database with data on the the level individual assignments
  
## Variables description
 
Important: 

  - **forum_attendance_tw** - Forum attendance during time window (start of assignment-first attempt). Binary.  
  - **forum_visits_tw** - the number of forum visits during time window. Counted.   
  - **CFA** - Correctness of the first attempt. Binary.

Less important:

  - **hse_user_id** - a student token;
  - **course_item_name** and **course_item_type_desc** - item token/name and item type;
  
## Hypothesis 1

### Model 1

Checking the variance of the CFA (correctness of the first attempt) variable 

```{r}
# m10 <- glm(data = assignments, factor(CFA) ~ 1, family = "binomial")
# saveRDS(m10, "data/m10.rds")
```

```{r, message= FALSE, warning=FALSE, echo=FALSE}
system('7z x data/m10.zip -odata/.')
m10 <- readRDS( "data/m10.rds")
```

```{r}
summary(m10)
```


```{r}
# m1 <- glmer(data = assignments, factor(CFA) ~ 1 + (1|hse_user_id) + (1|course_item_name), family = "binomial")
# saveRDS(m1, "data/m1.rds")
```

```{r, message= FALSE, warning=FALSE, echo=FALSE}
system('7z x data/m1.zip -odata/.')
m1 <- readRDS( "data/m1.rds")
```

```{r}
summary(m1)
```


```{r}
sd_proficieny <- data.frame(VarCorr(m1))[5][1, 1]
sd_difficulty <- data.frame(VarCorr(m1))[5][2, 1]
coef <- summary(m1)$coef[1] # intercept of m0

c_proficieny <- c(round(invlogit(coef - sd_proficieny), digits = 2), round(invlogit(coef + sd_proficieny), digits = 2))
c_difficulty <- c(round(invlogit(coef - sd_proficieny - sd_difficulty), digits = 2), round(invlogit(coef - sd_proficieny + sd_difficulty), digits = 2))

ll1 <- logLik(m10)*-2
ll2 <- logLik(m1)*-2
chi <- ll1[1] - ll2[1]
df <- 3-1
```

```{r, echo=FALSE}
cat(sprintf("Parameters of the Correctness of first attempt (CFA): intercept - %.2f, beta - %.2f.\n", round(invlogit(coef), digits = 2), coef))
cat(sprintf("Student standart deviation(proficiency) - %.2f.\n", sd_proficieny))
cat(sprintf("Assignment standart deviation (difficulty) - %.2f.\n", sd_difficulty))
cat(sprintf("-sd %.2f and +sd %.2f for student proficiency.\n", c_proficieny[1], c_proficieny[2] ))
cat(sprintf("-sd %.2f and +sd %.2f for assignment difficulty.\n", c_difficulty[1], c_difficulty[2]))
cat(sprintf("The results are significant if noncentral (q) chi-squared greater than chi \nchi-squared - %.2f, degrees of freedom  - %.2f, qchi-squared  - %.2f \n", chi, df, qchisq(p=.0001, df=df, lower.tail=FALSE)))
```

```{r}
anova(m10, m1)
```

```{r}
assignments  %>% filter(forum_visits_tw < 1000) %>%   ggplot(aes(x = forum_visits_tw, y = CFA, color = course_item_type_desc)) +
  geom_point()   +  

  geom_smooth(method = "glm", method.args=list(family="binomial")) +   
  ggtitle("Probability to pass the test from the first attempt\nand forum visits during first attempt") +
  labs(x = "Forum visists during first attempt", y = "Probability to pass test during first attempt", color = "Assignment type") +
  theme_bw() +  
  theme(plot.title = element_text(hjust = 0.5))
```



### Model 2

Checking the relationship between forum attendance and CFA with two random effects, on student proficience and course difficulty.

```{r}
# m2 <- glmer(data = assignments, factor(CFA) ~ 1 + forum_attendance_tw  + (1|hse_user_id) + (1|course_item_name), family = "binomial")
# saveRDS(m2, "data/m2.rds")
```

```{r}
system('7z x data/m2.zip -odata/.')
m2 <- readRDS("data/m2.rds")
summary(m2)
m20 <- summary(m2)
```

```{r, echo=FALSE}
cat(sprintf("Probability of CFA if a student didn't visit a forum before - %.2f.\nProbability of CFA if a student visited a forum at least once - %.2f\n\n",  invlogit(m20$coefficients[1])), invlogit(m20$coefficients[1] + m20$coefficients[2]))
```

#### Visualisation

```{r, warning=FALSE}
assignments  %>%  ggplot(., aes(factor(forum_attendance_tw), fill = factor(CFA))) +
  geom_bar(position = "dodge2") +
  labs(x = "Forum visited or not", fill = "First attempt") + theme_classic() +  
  scale_fill_discrete(labels=c("Not passed", "Passed")) +
  theme(plot.title = element_text(hjust = 0.5))


```
